Enhanced proximity effect in zigzag-shaped Majorana Josephson junctions.

### 2019-02-07 paper/snakemajoranas.tex
corrections with Tom

\documentclass[english, twocolumn, 10pt, aps, superscriptaddress, floatfix, prb, citeautoscript]{revtex4-1}
\pdfoutput=1
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{verbatim}
\usepackage{units}
\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{wasysym}
\usepackage{layouts}
\usepackage{siunitx}
\usepackage{bm}
\usepackage{xcolor}
\usepackage[colorlinks, citecolor={blue!50!black}, urlcolor={blue!50!black}, linkcolor={red!50!black}]{hyperref}
\usepackage{bookmark}
\usepackage{tabularx}
\usepackage{microtype}
\usepackage{babel}
\hypersetup{pdfauthor={Quantum Tinkerer},pdftitle={Increasing the topological gap by an order of magnitude through the geomerty for Majorana SNS junctions.}}

@@

\title{Increasing the topological gap by an order of magnitude through the geometry for Majorana SNS junctions.}

@@

\date{\today}
\begin{abstract}
We solve the problem of the appearance of a soft gap in high-density SNS Majorana junctions \cite{pientka2017topological}.
Trajectories that do not encounter the superconductor for a long time weaken the proximity effect in Josephson junctions.
A small induced gap then becomes an essential obstacle in creating Majorana states.
We show that a zigzag-shaped geometry eliminates these trajectories, at the same time allowing the robust creation of Majorana states with a gap increased by more than an order of magnitude.
\end{abstract}

@@

\comment{Majoranas are mostly made in hybrid NS structures.}
The current experimental effort ~\cite{mourik_signatures_2012,das_zero-bias_2012,deng_anomalous_2012,churchill_superconductor-nanowire_2013,zhang2018quantized} focuses on creating pairs of MBSs in hybrid normal-superconductor (NS) structures, where a long strip (or wire) of a semiconductor is interfaced with a superconductor.~\cite{lutchyn_majorana_2010,oreg_helical_2010}
Within the right parameter regime ($E_z^2>\mu^2+\Delta^2$ for a simple wire), the semiconductor becomes topological while MBSs appear at the ends of the segment.
Recently, a system with a slight modification has been proposed\cite{pientka2017topological}, where instead of having one superconductor, there are two superconductors, creating a superconductors-normal-superconductor (SNS) junction.
The additional superconductor adds an extra knob to adjust, the superconducting phase difference $\phi$, and this should make it easier to tune the system into the topological phase.
Recently, two groups~\cite{fornieri_evidence_2018,ren_topological_2018} have realized this system experimentally.
In practice, multiple problems arise making even the observation difficult, rendering the manipulation of Majoranas in the near future, improbable.

\comment{The gap is small because of long trajectories.}
One of the biggest challenges in creating stable Majoranas is the appearance of a soft gap; where the gap in the density of states significantly reduces for states with the momentum directed along the length of the strip.  % TODO: do we cite here?
From a semiclassical perspective, these momenta correspond to long paths through the semiconductor without interruption by the superconductor.
Additionally, these long trajectories have long flight times $\tau_f$ and equivalenly small Thouless energies $E_{\textrm{Th}}=\hbar / \tau_f$, resulting in a small gap.

\comment{Currently proposed workarounds are low density or disorder, but both have drawbacks.}
Currently, proposed workarounds are to work at low density \cite{nijholt2015orbital} or to have disorder~\cite{haim_double-edge_2018}, but both have drawbacks.

@@

\comment{We consider a two-dimensional system with zigzag.}
We consider a Josephson junction (JJ)(Fig.~\ref{fig:setup}) consisting of a two-dimensional strip of semiconductor, with superconductors on both sides.
We modulate the shape of the normal region, which can be either zigzag as depicted, or a more smooth sinusoidal-like shape.
Similar to the conventional straight system \cite{pientka2017topological}, a magnetic field $B_x$ perpendicular to the junction is applied.
We model the system with a Bogoliubov-de Gennes Hamiltonian (BdG) Eq.~\eqref{eq:hamiltonian}; the normal part has a linear Rashba spin-orbit coupling term characterized by $\alpha$ and a Zeeman field with $E_z=\frac{1}{2} \mu_B g B_x$.
The superconductor has a superconducting coupling term $\Delta$ with the pair of superconductors at an opposite superconducting phase $\exp{\pm i \phi/2}$.
Both the normal part and the superconducters have a kinetic term and chemical potential $\mu$.
The complete Hamiltonian has the form
\begin{equation}
\begin{split}
    H = & \left[\frac{\hbar^2\left(\kx^2 + \ky^2\right)}{2\meff} - \mu\right] \tau_z \\
        & + E_\text{z}(x,y) \sigma_x \\
        & + \alpha(x,y) \left( \ky \sigma_x - \kx \sigma_y \right) \tau_z
          + \Delta(x,y) \tau_y.
\end{split}
\label{eq:hamiltonian}
\end{equation}
It acts on the spinor wave function $\Psi={\left(\psi_{e\uparrow},\psi_{e\downarrow},\psi_{\textrm{h}\downarrow},-\psi_{\textrm{h}\uparrow}\right)}^{T}$, where $\psi_e$, $\psi_\textrm{h}$ are its electron and hole components, and $\psi_\uparrow$, $\psi_\downarrow$ are the spin-up and spin-down components.
The Pauli matrices $\sigma_{i}$ act on the spin degree of freedom and $\tau_{i}$ act on the electron-hole degree of freedom.
For the zigzag geometry, we first consider a sawtooth-like pattern where $z_x$ is periodicity, $z_y$ the amplitude and $W$ the width of the junction.
Later in this manuscript, we show that the exact details of the shape are unimportant and consider different zigzag geometries also parametrized by $z_x, \; z_y, \; W$.

Unless noted differently, the Hamiltonian parameters are $\alpha=\SI{20}{\meV \nm}$, $g=26$, $\meff=\SI{0.02}{\electronmass}$, $\mu=\SI{10}{\meV}$, $B=\SI{1}{T}$, $\phi=\pi$, and $\Delta=\SI{1}{\meV}$; and the geometry parameters are $W=\SI{200}{\nm}$, the period of the zigzag $z_x=\SI{1300}{\nm}$, the discretization contant $a=\SI{10}{\nm}$, and the lengths of the superconductors $L_\textrm{SC}=\SI{200}{\nm}$.

@@

\comment{We discretize the Hamiltonian and simulate it with Kwant.}
We discretize our continuum Hamiltonian [Eq.~\eqref{eq:hamiltonian}] on a square grid and implement a tight-binding model using Kwant~\cite{groth_kwant:_2014}.
The resulting Hamiltonians are large because instead of having a single slice of sites that is translationally invariant in the $x$-direction, we create a long supercell to make the structure periodic.
To speed up these expensive simulations we employ an adaptive sampling strategy, using the Python package Adaptive.~\cite{adaptive}
The entire source code is available in the Supplemental Material~\cite{supp}.  % TODO: update the bib when the data is uploaded
The resulting raw data are available in Ref.~\onlinecite{data}.  % TODO: update the bib when the data is uploaded

@@

\comment{In a straight system, the Majoranas are very poorly localized.}
For the ribbon system [Fig.~\ref{fig:wavefunctions}(a)], we see that the decay of the density is long compared to the system size.
The wavefunction extends to the center of the system, resulting in overlapping Majoranas and a non-zero energy of the state.
Further, the energy of the Majorana state $E_\textrm{M}$ in the straight system is badly septated from the next lowest lying eigenstate $E_1$.
Altogether, the increase in the topological gap, but also the before mentioned reduction of the group velocity, result in poorly localized Majoranas and minimal topological protection against perturbations.  % TODO: probably put the eq for coherence length here?

>>>>

\documentclass[english, twocolumn, 10pt, aps, superscriptaddress, floatfix, prb, citeautoscript]{revtex4-1}
\pdfoutput=1
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{verbatim}
\usepackage{units}
\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{wasysym}
\usepackage{layouts}
\usepackage{siunitx}
\usepackage{bm}
\usepackage{xcolor}
\usepackage[colorlinks, citecolor={blue!50!black}, urlcolor={blue!50!black}, linkcolor={red!50!black}]{hyperref}
\usepackage{bookmark}
\usepackage{tabularx}
\usepackage{microtype}
\usepackage{babel}
\hypersetup{pdfauthor={Quantum Tinkerer},pdftitle={Enhanced proximity effect in zigzag-shaped Majorana Josephson junctions.}}

@@

\title{Enhanced proximity effect in zigzag-shaped Majorana Josephson junctions.}

@@

\date{\today}
\begin{abstract}
We solve the problem of the appearance of a soft gap in high-density SNS Majorana junctions \cite{pientka2017topological}.
Trajectories that do not encounter the superconductor for a long time weaken the proximity effect in Josephson junctions.
A small induced gap then becomes an essential obstacle in creating Majorana states.
We show that a zigzag-shaped geometry eliminates these trajectories, at the same time allowing the robust creation of Majorana states with a gap increased by more than an order of magnitude and a significantly reduced Majorana decay length.
\end{abstract}

@@

\comment{Majoranas are mostly made in hybrid NS structures.}
The current experimental effort ~\cite{mourik_signatures_2012,das_zero-bias_2012,deng_anomalous_2012,churchill_superconductor-nanowire_2013,zhang2018quantized} focuses on creating pairs of MBSs in hybrid normal-superconductor (NS) structures, where a long strip (or wire) of a semiconductor is interfaced with a superconductor.~\cite{lutchyn_majorana_2010,oreg_helical_2010}
Within the right parameter regime ($E_z^2>\mu^2+\Delta^2$ for a simple wire), the semiconductor becomes topological while MBSs appear at the ends of the segment.

\comment{A alternative has been proposed (JJ).}
Recently, a system with a slight modification has been proposed\cite{pientka2017topological}, where instead of having one superconductor, there are two superconductors, creating a superconductors-normal-superconductor (SNS) junction.
The additional superconductor adds an extra knob to adjust, the superconducting phase difference $\phi$, and this should make it easier to tune the system into the topological phase.
Recently, two groups~\cite{fornieri_evidence_2018,ren_topological_2018} have realized this system experimentally.
In practice, multiple problems arise making even the observation difficult, rendering the manipulation of Majoranas in the near future, improbable.

\comment{The gap is small because of long trajectories.}
One of the biggest challenges in creating stable Majoranas is the appearance of a soft gap~\cite{ren_topological_2018}; where the gap in the density of states significantly reduces for states with the momentum directed along the length of the strip.
From a semiclassical perspective, these momenta correspond to long paths through the semiconductor without interruption by the superconductor.
Additionally, these long trajectories have long flight times $\tau_f$ and equivalenly small Thouless energies $E_{\textrm{Th}}=\hbar / \tau_f$, resulting in a small gap.

\comment{Currently proposed workarounds are low density or disorder, but both have drawbacks.}
Currently, proposed workarounds are to work at low density \cite{nijholt2015orbital} or to have disorder~\cite{haim_double-edge_2018}, but both have drawbacks.  % TODO: mention drawbacks?

@@

\comment{We consider a two-dimensional system with zigzag.}
We consider a Josephson junction (JJ)(Fig.~\ref{fig:setup}) consisting of a two-dimensional strip of semiconductor, with superconductors on both sides.
We modulate the shape of the normal region, which can be either zigzag as depicted, or a more smooth sinusoidal-like shape.
Similar to the conventional straight system \cite{pientka2017topological}, a magnetic field $B_x$ perpendicular to the junction is applied.
We model the system with a Bogoliubov-de Gennes Hamiltonian (BdG) Eq.~\eqref{eq:hamiltonian}; the normal part has a linear Rashba spin-orbit coupling term characterized by $\alpha$ and a Zeeman field with $E_z=\frac{1}{2} \mu_B g B_x$.
The superconductor has a superconducting coupling term $\Delta$ with the pair of superconductors at an opposite superconducting phase $\exp{\pm i \phi/2}$.
Both the normal part and the superconducters have a kinetic term and chemical potential $\mu$.
The complete Hamiltonian has the form
\begin{equation}
\begin{split}
    H = & \left[\frac{\hbar^2\left(\kx^2 + \ky^2\right)}{2\meff} - \mu\right] \tau_z \\
        & + E_\text{z}(x,y) \sigma_x \\
        & + \alpha(x,y) \left( \ky \sigma_x - \kx \sigma_y \right) \tau_z
          + \Delta(x,y) \tau_y.
\end{split}
\label{eq:hamiltonian}
\end{equation}
It acts on the spinor wave function $\Psi={\left(\psi_{e\uparrow},\psi_{e\downarrow},\psi_{\textrm{h}\downarrow},-\psi_{\textrm{h}\uparrow}\right)}^{T}$, where $\psi_e$, $\psi_\textrm{h}$ are its electron and hole components, and $\psi_\uparrow$, $\psi_\downarrow$ are the spin-up and spin-down components.
The Pauli matrices $\sigma_{i}$ act on the spin degree of freedom and $\tau_{i}$ act on the electron-hole degree of freedom.
For the zigzag geometry, we first consider a sawtooth-like pattern where $z_x$ is periodicity, $z_y$ the amplitude and $W$ the width of the junction.
The spatial dependence of $\Delta$, $E_z$, and $\alpha$ is explained in text above.
Later in this manuscript, we show that the exact details of the shape are unimportant and consider different zigzag geometries also parametrized by $z_x, \; z_y, \; W$.

Unless noted differently, the Hamiltonian parameters are $\alpha=\SI{20}{\meV \nm}$, $g=26$, $\meff=\SI{0.02}{\electronmass}$, $\mu=\SI{10}{\meV}$, $B=\SI{1}{T}$, $\phi=\pi$, and $\Delta=\SI{1}{\meV}$; and the geometry parameters are $W=\SI{200}{\nm}$, the period of the zigzag $z_x=\SI{1300}{\nm}$, the discretization contant $a=\SI{10}{\nm}$, and the lengths of the superconductors $L_\textrm{SC}=\SI{200}{\nm}$.  % TODO: Tom: L_sc and Delta not compatible?

@@

\comment{We discretize the Hamiltonian and simulate it with Kwant.}
We discretize our continuum Hamiltonian [Eq.~\eqref{eq:hamiltonian}] on a square grid and implement a tight-binding model using Kwant~\cite{groth_kwant:_2014}.
The resulting Hamiltonians are large because instead of having a single slice of sites that is translationally invariant in the $x$-direction, we create a long 2D supercell to make the structure periodic.
To speed up these expensive simulations we employ an adaptive sampling strategy, using the Python package Adaptive.~\cite{adaptive}
The entire source code is available in the Supplemental Material~\cite{supp}.  % TODO: update the bib when the data is uploaded
The resulting raw data are available in Ref.~\onlinecite{data}.  % TODO: update the bib when the data is uploaded

@@

\comment{In a straight system, the Majoranas are very poorly localized.}
For the ribbon system [Fig.~\ref{fig:wavefunctions}(a)], we see that the decay of the density is long compared to the system size.
The wavefunction extends to the center of the system, resulting in overlapping Majoranas and a non-zero energy of the state.
Further, the energy of the Majorana state $E_\textrm{M}$ in the straight system is badly separated from the next lowest lying eigenstate $E_1$.
Altogether, the increase in the topological gap, but also the before mentioned reduction of the group velocity, result in poorly localized Majoranas and minimal topological protection against perturbations.  % TODO: probably put the eq for coherence length here?
